Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-01-18 paper.tex
Finish Francois' edits, and clean up previous edits

The matter potential along this test ray is from an analytic wind
model qualitatively consistent with the densities in the
simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-2}$
density field of a spherical steady-state wind with a constant asymptotic
velocity.
The velocity and density fields of a spherical steady-state wind
obey the continuity equation $\rho(r)v(r)r^2=\rho_1 v_1 r_1^2$,
with $\rho_1$ and $v_1$ the density and velocity measured at some fiducial
radius $r_1$.
For velocity field we choose a phenomenological wind model used in the
oscillation literature\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
\begin{equation}
  v(r) = v_1 \left(1-\frac{R}{r_1}\right)^{-\beta}
  \left(1-\frac{R}{r}\right)^{\beta},
\end{equation}
with $R$ the wind launch radius and $\beta$ an acceleration parameter.
This yields the following density field
\begin{equation}
  \rho(r)=\rho_1 C\left[\frac{r_1}{R}\right]
  \left(1-\frac{R}{r}\right)^{-\beta}\left(\frac{R}{r}\right)^2,
\end{equation}
with $C[a]=a^{2-\beta}(a-1)^\beta$.
We use the parameters $R=10$~km, $R_1=50$~km, $\beta=2$,
and $\rho_1=10^7\,{\rm g}\,{\rm cm}^{-3}$.
Additionally we impose a density cap of
$\rho_{\rm max}=10^{14}\,{\rm g}\,{\rm cm}^{-3}$ inside the radius $r_0=5$~km,
and smoothly interpolate densities between $r_0$ and $r_1$ with a cubic polynomial,
enforcing $C^0$ and $C^1$ continuity at the transitions.
The rest density from this model is plotted in Fig.~\ref{fig:test_traj_density}.
To translate this density field to a matter potential, we assume a constant
electron fraction of $Y_e=0.5$.

@@

In fact a spherical steady-state wind model, though providing an adequate
backdrop for the qualitative study presented in this section,
is less than ideal for this post-merger configuration,
most obviously because in the 10~ms since merger,
ejecta with the greatest velocities around $0.3c$
will have reached no further than $r_{\rm max} \sim 10^8$~cm.
Additionally, the true radial profile of the ejecta from this merger
(which our computational model does not follow)
will have many more features inside this radius, including shock jumps.
This model density profile is adequate as a backdrop to our study here, however,
since we expect the remnant to present similar neutrino emission over a thermal
timescale of a few tens of milliseconds\footnote{cite Nouri et al 2018}
while the matter field propagates out to larger radii.

@@

We use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum_2}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

>>>>

The matter potential along this test ray is from an analytic wind
model qualitatively consistent with the densities in the
simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-2}$
density field of a spherical steady-state wind with a constant asymptotic
velocity.
The density and velocity fields of a spherical steady-state wind
obey the continuity equation $\rho(r)v(r)r^2=\rho_1 v_1 r_1^2$,
with $\rho_1$ and $v_1$ the density and velocity measured at some fiducial
radius $r_1$.
For velocity field we choose a phenomenological wind model used in the
oscillation literature \footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
\begin{equation}
  v(r) = v_1 \left(1-\frac{R}{r_1}\right)^{-\beta}
  \left(1-\frac{R}{r}\right)^{\beta},
\end{equation}
with $R$ the wind launch radius and $\beta$ an acceleration parameter.
This yields the following density field
\begin{equation}
  \rho(r)=\rho_1 C\left[\frac{r_1}{R}\right]
  \left(1-\frac{R}{r}\right)^{-\beta}\left(\frac{R}{r}\right)^2,
\end{equation}
with $C[a]=a^{2-\beta}(a-1)^\beta$.
We use the parameters $R=10$~km, $r_1=50$~km, $\beta=2$,
and $\rho_1=10^7\,{\rm g}\,{\rm cm}^{-3}$.
Additionally we impose a density cap of
$\rho_{\rm max}=10^{14}\,{\rm g}\,{\rm cm}^{-3}$ inside the radius $r_0=5$~km,
and smoothly interpolate densities between $r_0$ and $r_1$ with a cubic polynomial,
enforcing $C^0$ and $C^1$ continuity at the transitions.
The rest density from this model is plotted in Fig.~\ref{fig:test_traj_density}.
To translate this density field to a matter potential, we assume a constant
electron fraction of $Y_e=0.5$, roughly consistent with the composition of the
matter in the disk's funnel \cite[Fig.~8]{fouc2016-m1_evolve_n}.

@@

In fact a spherical steady-state wind model, though providing an adequate
backdrop for the qualitative study presented in this section,
is less than ideal for this post-merger configuration,
most obviously because in the 10~ms since merger,
ejecta with the greatest velocities around $0.3c$
will have reached no further than $r_{\rm max} \sim 10^8$~cm.
Additionally, the true radial profile of the ejecta from this merger
(which our computational model does not follow)
will have many more features inside this radius, including shock jumps.
This model density profile is adequate as a backdrop to our study here, however,
since we expect the remnant to present similar neutrino emission over a thermal
timescale of a few tens of milliseconds \footnote{cite Nouri et al 2018}
while the matter field propagates out to larger radii.

@@

We use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum_2}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is the lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$.
